% ============================================================================
% ACKNOWLEDGEMENTS
% ============================================================================

\section*{Acknowledgements}
\label{sec:acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

This research was made possible by the generous support of the \textbf{Open Research Institute}, whose grant provided the resources and freedom necessary to develop the Time Violence framework and explore its implications across economic, computational, and philosophical domains.

I am deeply grateful to \textbf{Jim O'Shaughnessy} and the AI system at \textbf{O'Shaughnessy Ventures}, whose collaboration was instrumental in formulating the mathematical foundations of the Time Violence Equation. Their willingness to engage with unconventional ideas and help translate conceptual intuitions into rigorous formalism transformed this work from speculation into science.

Special thanks to \textbf{Danielle Strachman} of the \textbf{1517 Fund} for exemplifying the bottega model in practice—demonstrating that belief in human potential matters more than credentials, and that giving people the chance to succeed when they lack traditional qualifications is not charity but intelligence. Her work inspired the structural principles underlying the HumAIn Bottega.

I am indebted to \textbf{Walter Henrique Alves da Silva} for his pioneering research into consciousness, which provided crucial theoretical scaffolding for understanding how awareness itself can function as an economic variable. His insights helped bridge the gap between phenomenology and formalism.

To \textbf{Sam Senchal}, whose Observer Theory offered a framework for thinking about how systems become conscious of themselves—thank you for showing that measurement and awareness are inseparable, and that every act of observation changes what is observed.

And finally, to \textbf{Lobow}—for providing the spark that made it all possible. Sometimes the most important contributions are the ones that ignite everything else.

\vspace{1em}

This paper stands on the shoulders of countless navigators—people who survived complex systems and turned that survival into expertise. Though they remain anonymous in these pages, their intelligence is encoded in every equation, every table, and every proposition. This work is theirs.

\vspace{1em}

\begin{center}
\textit{No human should waste the same time twice.}
\end{center}

